Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results

### 2014-10-31 workshop-paper/paper.tex
Experiments draft

\begin{table*}
%% \addtolength{\tabcolsep}{-2pt}
{\scriptsize
\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
%% table header
\hline\\[-3ex]
\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{BrFS}& \multicolumn{3}{c|}{IW1}& \multicolumn{3}{c|}{BFS}& \multicolumn{3}{c}{UCT} \\[.2ex]
\hline\\[-3ex]
%% table body
\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{Dev}  & \multicolumn{1}{c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev} \\[.2ex]
\hline\\[-3ex]
asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
\end{tabular}
}

 \caption{\small
   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
   seeds 0 to 4.  \textit{T} is the action selection average
   computation time, \textit{S} is the average score, and
   \textit{Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
\label{performance}

\end{table*}

** Results, cold and detached, same format as those reported in jair paper, possibly with
std deviations and times.

>>>>

\input{performance_table}

The algorithms IW1 and BFS are tested over 9 games selected from those
considered in the training set in \cite{bellemare:jair}, and those
considered in \cite{deep-mind-atari}. Each game has 5 runs, each with
a different random seed $0,\ldots,4$.  We compare the performance of
IW1 and BFS with UCT and BRFS. We also include results for a baseline random
agent. We use the same parameters in the literature, all algorithms use a discount factor of 0.999, and each action
simulates 5 frames. Each algorithm has a budget on the number of nodes
it can generate per lookahead. BRFS, IW1 and BFS have a maximum budget
of 100k. For UCT we use the same parameters as in
\cite{bellemare:jair}, 0.1 exploration rate, and 500 rollouts of a maximum depth of 300 steps,
which gives a maximum budget of 150k nodes. We tested UCT with a fix
budget of 100k and max depth 300, but resulted in practically the same
results. All algorithms reuse the tree built between lookaheads by
setting the new root of the tree to the node of the selected action,
and deleting the other branches. Note that this may have a different
effect on algorithms were all branches have the same number of
generated nodes, ex. BRFS, than those that may focus on different
areas of the tree, ex. IW1, BFS, UCT.

On Table~\ref{performance}, we can observe the performance of each
algorithm on terms of average score with standard deviation, and
average time per lookahead in seconds. The Random agent don't include
average time as it just chooses an action randomly. BRFS outperform
the Random agent in 7 out of 9 domains. BRFS rarely finds any reward,
turning itself mostly into a random agent, but exploring 100k nodes up
to an average depth of 4. When BRFS is turned int IW1 using novelty to
explore the search space, the jump on performance is important. IW1 in
most of the cases does not completely use its budget, and is able to
achieve depths of 70 to 260 depending on the game, which first
explains how IW1 is able to achieve a performance that is closer to
UCT than to BRFS. In 4 out of 9 tested games it is the best
performer. Remarkably, is the first domain independent solver that
shows good performance on Freeway, where rewards are very sparse and
deep in the search tree. BFS outperforms IW1 and UCT in 4 and 2
domains respectively.  Both IW1 and BFS performance in terms of score
is close to UCT, but they are some times one order of magnitud
faster. The speedup is due to the bounded space in IW1, and to the
fact that in some games, both IW1 and BFS are able to reuse an
important part of their trees between lookaheads.

The novelty of states in IW1 and BFS may depend on the order that
nodes are generated. Therefore, every time a node is expanded, we
randomized the order in which nodes get genreated. Randomizing the
order of expansion is also applied to BRFS, but has no impact on its
performance.

IW2 is not included in the table due to its low performance, which is
closer to BRFS than to the other algorithms. This is mainly due to the
huge number of states that are novel in this algorith, thus,
achieveing insufficient depths to gather some reward.
